---
title: "Movie Recommender"
author: "H.Gryk"
date: "2020-09-15"
categories: 
  -Shiny
  -Recommender
tags:
  -Movie Recommender
  -Shiny App
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE}
library(utils)
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```

I was recently inspired to create a movie recommender when a family member started hosting a virtual movie night.  Given the variety of movie preferences of the invited group, I was motivated to create recommendations based on multiple genres and movies. 

The data used for this task is from [movielens](https://grouplens.org/datasets/movielens/).  The data available from movielens includes links, movies, ratings, and tags, however, only some of this data will be utilized herin.  


```{r echo=FALSE}
fileLoc <-"/Users/Hailey/Documents/GitHub/projects/MoviePredictor/ml-latest-small.zip"
#unzip(fileLoc)

movies <- read.csv("ml-latest-small/movies.csv")
ratings <- read.csv( "ml-latest-small/ratings.csv")
# tags <- read.csv("ml-25m/tags.csv")
# mv_tags <- read.csv("ml-25m/genome-scores.csv")
# mv_tags_desc <- read.csv("ml-25m/genome-tags.csv")
```



```{r}
# preprocessing data

genres <- as.data.frame(movies$genres, stringsAsFactors = FALSE)
genres2 <- as.data.frame(tstrsplit( genres[,1], '[|]',
                                    type.convert=TRUE),
                                    stringsAsFactors=FALSE)

colnames(genres2) <- c(1:10)
genre_list <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western") #list of all the genres

genre_matrix <-matrix(0, nrow(movies)+1, length(genre_list)) #empty matrix, no movies + 1, no genres
genre_matrix[1,] <- genre_list # populate genres in first row
colnames(genre_matrix) <- genre_list # set column names with genre list

#iterate through matrix
for (i in 1:nrow(genres2)) {
  for (c in 1:ncol(genres2)) {
    genmat_col = which(genre_matrix[1,] == genres2[i,c])
    genre_matrix[i+1,genmat_col] <- 1
  }
}

#convert into dataframe
genre_matrix2 <- as.data.frame(genre_matrix[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (c in 1:ncol(genre_matrix2)) {
  genre_matrix2[,c] <- as.integer(genre_matrix2[,c])
} #convert from characters to integers
```

```{r}
#Create a matrix to search for a movie by genre:
years <- as.data.frame(movies$title, stringsAsFactors=FALSE)
library(data.table)
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
years <- as.data.frame(substr(substrRight(substrRight(years$`movies$title`, 6),5),1,4))

search_matrix <- cbind(movies[,1], substr(movies[,2],1,nchar(movies[,2])-6), years, genre_matrix2)
colnames(search_matrix) <- c("movieId", "title", "year", genre_list)

write.csv(search_matrix, "search.csv")
search_matrix <- read.csv("search.csv", stringsAsFactors=FALSE)

# at this point we can search for movies with the following:
#subset(search_matrix, Crime ==1 & year == 1990)$title
```

```{r}
#ratings matrix -1 for less than 3, 1 for more than 3

binaryratings <- ratings

for (i in 1:nrow(binaryratings)){
  if (binaryratings[i,3] > 3){
    binaryratings[i,3] <- 1
  }
  else{
    binaryratings[i,3] <- -1
  }
}

# convert binaryratings matrix to the correct format:
binaryratings2 <- dcast(binaryratings, movieId~userId, value.var = "rating", na.rm=FALSE)

for (i in 1:ncol(binaryratings2)){
  binaryratings2[which(is.na(binaryratings2[,i]) == TRUE),i] <- 0
}
binaryratings2 = binaryratings2[,-1] #remove movieIds col. Rows are movieIds, cols are userIds


#Remove rows that are not rated from movies dataset
movieIds <- length(unique(movies$movieId)) #10329
ratingmovieIds <- length(unique(ratings$movieId)) #10325
movies2 <- movies[-which((movies$movieId %in% ratings$movieId) == FALSE),]
rownames(movies2) <- NULL

#Remove rows that are not rated from genre_matrix2
genre_matrix3 <- genre_matrix2[-which((movies$movieId %in% ratings$movieId) == FALSE),]
rownames(genre_matrix3) <- NULL


```

```{r}
# find dot product of genre matrix and user ratings
#Calculate dot product for User Profiles
result = matrix(0,length(genre_list), max(ratings$userId)) # here, no of users/raters, no of genres
for (c in 1:ncol(binaryratings2)){
  for (i in 1:ncol(genre_matrix3)){
    result[i,c] <- sum((genre_matrix3[,i]) * (binaryratings2[,c])) #ratings per genre
  }
}

#Convert to Binary scale
for (c in 1:ncol(result)){
  for (i in 1:nrow(result)){
    if (result[i,c] < 0){
      result[i,c] <- 0
    }
    else {
      result[i,c] <- 1
    }
  }
}

```

```{r}
# get movies that are closest in similarity to the genres & user ratings (i.e. the representation of user's preference)
# use Jaccard distance to measure similarities for user based filtering

#Create ratings matrix. Rows = userId, Columns = movieId
ratingmat <- dcast(ratings, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingmat <- as.matrix(ratingmat[,-1]) #remove userIds

# Similarity Calculation with Cosine Similarity nearest neighbors of 30
#Convert rating matrix into a recommenderlab sparse matrix
ratingmat <- as(ratingmat, "realRatingMatrix")

# Determine how similar the first four users are with each other
# create similarity matrix
similarity_users <- similarity(ratingmat[1:4, ], 
                               method = "cosine", 
                               which = "users")
as.matrix(similarity_users)
image(as.matrix(similarity_users), main = "User similarity")

# compute similarity between
# the first four movies
similarity_items <- similarity(ratingmat[, 1:4], method =
                                 "cosine", which = "items")
as.matrix(similarity_items)
image(as.matrix(similarity_items), main = "Item similarity")


```

```{r}
# Exploring values of ratings:
vector_ratings <- as.vector(ratingmat@data)
unique(vector_ratings) # what are unique values of ratings

table_ratings <- table(vector_ratings) # what is the count of each rating value
table_ratings

# Visualize the rating:
vector_ratings <- vector_ratings[vector_ratings != 0] # rating == 0 are NA values
vector_ratings <- factor(vector_ratings)

qplot(vector_ratings) + 
  ggtitle("Distribution of the ratings")
```

```{r}
# Exploring viewings of movies:
views_per_movie <- colCounts(ratingmat) # count views for each movie

table_views <- data.frame(movie = names(views_per_movie),
                          views = views_per_movie) # create dataframe of views
table_views <- table_views[order(table_views$views, 
                                 decreasing = TRUE), ] # sort by number of views

ggplot(table_views[1:6, ], aes(x = movie, y = views)) +
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_discrete(labels=subset(movies2, movies2$movieId == table_views$movie)$title) +
  ggtitle("Number of views of the top movies")

#Visualizing the matrix:
image(ratingmat, main = "Heatmap of the rating matrix") # hard to read-too many dimensions
image(ratingmat[1:10, 1:15], main = "Heatmap of the first rows and columns")
image(ratingmat[rowCounts(ratingmat) > quantile(rowCounts(ratingmat), 0.99),
                 colCounts(ratingmat) > quantile(colCounts(ratingmat), 0.99)], 
      main = "Heatmap of the top users and movies")
```


      
```{r normalzing, eval=FALSE, include=FALSE}
#Normalize the data
ratingmat_norm <- normalize(ratingmat)
image(ratingmat_norm[rowCounts(ratingmat_norm) > quantile(rowCounts(ratingmat_norm), 0.99),
                colCounts(ratingmat_norm) > quantile(colCounts(ratingmat_norm), 0.99)], 
      main = "Heatmap of the top users and movies")
```

```{r}
#Create UBFC Recommender Model. UBCF stands for User-Based Collaborative Filtering
recommender_model <- Recommender(ratingmat_norm, 
                                 method = "UBCF", 
                                 param=list(method="Cosine",nn=30))

model_details <- getModel(recommender_model)
model_details$data

recom <- predict(recommender_model, 
                 ratingmat[1], 
                 n=5) #Obtain top recommendations for 1st user in dataset

recom

recom_list <- as(recom, 
                 "list") #convert recommenderlab object to readable list

```

```{r}
#Obtain recommendations
recom_result <- matrix(0,5)
for (i in 1:length(recom_result)){
  recom_result[i] <- as.character(subset(movies, 
                                         movies$movieId == as.integer(recom_list[[1]][i]))$title)
}

```








